---
title: "Homework 6, Lab 6, Econ B2000"
author: 'Nicole Kerrison & Christopher Tinevra'
date: "11/9/2020"
output: github_document
---

#### Lab Results for Lab 6 "Logic and Probit Models" 

```{r}
load("~/CCNY/Graduate School/Fall 2020/Econometrics/ACS Data for R/acs2017_ny_data.RData")
suppressMessages(attach(acs2017_ny))
```

```{r}
acs2017_ny$EMPSTAT <- as.factor(acs2017_ny$EMPSTAT)
levels(acs2017_ny$EMPSTAT) <- c("NA","Employed","Unemployed","Not in LF")

acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
levels(acs2017_ny$LABFORCE) <- c("NA","Not in LF","in LF")

acs2017_ny$MARST <- as.factor(acs2017_ny$MARST)
levels(acs2017_ny$MARST) <- c("married spouse present","married spouse absent","separated","divorced","widowed","never married")

table(acs2017_ny$EMPSTAT,acs2017_ny$LABFORCE)

acs2017_ny$age_bands <- cut(acs2017_ny$AGE,breaks=c(0,25,35,45,55,65))
table(acs2017_ny$age_bands,acs2017_ny$LABFORCE)

```

Tables demonstrates the total amount of people from the dataset that are part of the labor force. Additionally, the people are divided into age categories ranging from 25 to 65 years old and the amount of people that are part of the labor force by age group. 

```{r}
use_varb <- (acs2017_ny$AGE >25) & (acs2017_ny$AGE <= 55)
acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
dat_use <- subset(acs2017_ny,use_varb)
dat_use$LABFORCE <- droplevels(dat_use$LABFORCE) 
model_logit <- glm(LABFORCE ~ AGE ,family = binomial, data = dat_use)
model_probit <- glm(LABFORCE ~ AGE ,family = binomial (link = 'probit'), data = dat_use)

summary(dat_use$LABFORCE)
suppressMessages(require(stargazer))
stargazer(model_logit, model_probit, type = "text")

NNobs <- length(dat_use$LABFORCE)
set.seed(12345) 
graph_obs <- (runif(NNobs) < 0.8)
dat_graph <-subset(dat_use,graph_obs)  
 plot(LABFORCE ~ AGE, pch = 16, ylim = c(0,1), data = dat_graph, main = "Labor Force", xlab = "Age", ylab = "Labor Force Status", col = c("blue","gray"))
```

The following lab model illustrates a simple probit and logit non linear regression where the dependent variable is the labor force participation and the indenpent variable is age. 

```{r}
model_logit1 <- glm(LABFORCE ~ AGE + I(AGE^2) + female + AfAm + Asian + race_oth + Hispanic+ educ_hs + educ_somecoll + educ_college + educ_advdeg + MARST, family = binomial, data = dat_use)

model_probit1 <- glm(LABFORCE ~ AGE + I(AGE^2) + female + AfAm + Asian + race_oth + Hispanic+ educ_hs + educ_somecoll + educ_college + educ_advdeg + MARST, family = binomial (link = 'probit'), data = dat_use)

suppressMessages(require(stargazer))
stargazer(model_logit1, model_probit1, type = "text")
detach()
```

Model 1 illustrates a probit and logit non linear regression with multiple independent variables. These independent variables may have an effect on whether a person participates in the labor force or not.

Using the coefficient estimates from logic and probit model, we can predict various probabilities while using independent variables and keeping labor force as the dependent variable.

For example, using the logic model we can predict the probability of a 35 year old female with an advanced degree within the labor force considering the independent variables as AGE, AGE^2, FEMALE, and EDUC_ADV.

The following functions provides the probability prediction using the relevant coefficients and setting other independent variables to zero since its not applicable to the analysis. 

```{r}
1/(1+(exp(-(0.602 + (0.017)*(35) + (-0.0003)*(35^2) + (-0.684*1) + (2.368*1)))))
```
For the logic model, the predicted probability for a 35 year old female with an advanced degree to be part of the labor force is 92.51%

Another example, using the probit model we can predict the probability of a 35 year old female with an advanced degree within the labor force considering the independent variables as AGE, AGE^2, FEMALE, and EDUC_ADV.

```{r}
pnorm(0.602 + (0.017)*(35) + (-0.0003)*(35^2) + (-0.684*1) + (2.368*1))
```
For the probit model, the predicted probability for a 35 year old female with an advanced degree to be part of the labor force is 99.40%

When comparing the probabilities calculated from the probit and logit model, the difference is 6.89 percentage points. The probit model shows a stronger predictability of determining whether a female with a advanced degree is part of the labor force.

Furthermore, the logic and probit models can be further expanded by adding additional independent variables into the Generalized non-linear Models while maintaining Labor force as the dependent variable. In this case, we will be taking a look at the citizen status as the independent variable. 

```{r}
suppressMessages(attach(acs2017_ny))

acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
levels(acs2017_ny$LABFORCE) <- c("NA","Not in LF","in LF")

acs2017_ny$MARST <- as.factor(acs2017_ny$MARST)
levels(acs2017_ny$MARST) <- c("married spouse present","married spouse absent","separated","divorced","widowed","never married")

acs2017_ny$CITIZEN <- as.factor(acs2017_ny$CITIZEN)
levels(acs2017_ny$CITIZEN) <- c("Born abroad of American parents", "Naturalized citizen", "Not a citizen", "Not a citizen, but has received first papers", "Foreign born, citizenship status not reported")
```

```{r}
use_varb1 <- (acs2017_ny$AGE >25) & (acs2017_ny$AGE <= 55)

dat_use2 <- subset(acs2017_ny, use_varb1)

dat_use2$LABFORCE <- droplevels(dat_use2$LABFORCE)
dat_use2$MARST <- droplevels(dat_use2$MARST)
dat_use2$CITIZEN <- droplevels(dat_use2$CITIZEN)

model_logit2 <- glm(LABFORCE ~ AGE + I(AGE^2) + female + AfAm + Asian + race_oth + Hispanic + educ_hs + educ_somecoll + educ_college + educ_advdeg + MARST +CITIZEN, family = binomial, data = dat_use2)

model_probit2 <- glm(LABFORCE ~ AGE + I(AGE^2) + female + AfAm + Asian + race_oth + Hispanic + educ_hs + educ_somecoll + educ_college + educ_advdeg + MARST+ CITIZEN, family = binomial (link = 'probit'), data = dat_use2)

suppressMessages(require(stargazer))
stargazer(model_logit2, model_probit2, type = "text")

detach()
```

Model 2 illustrates a probit and logit non linear regression with multiple independent variables but this time taking into consideration citizen status. Using the coefficient estimates from logic and probit model 2, we can predict probabilities whether naturalize citizens and non citizens participate in the labor force.

For the logic model we can predict the probability of a 35 year old female naturalized citizen and non citizen with an college degree participates in the labor force while considering the independent variables as AGE, AGE^2, FEMALE, EDUC_COLLEGE, CITIZENNaturalized citizen and CITIZENNot a citizen. 

The following functions provides the probability prediction using the relevant coefficients and setting other independent variables to zero since its not applicable to the analysis. 

```{r}
1/(1+(exp(-(0.707 + (0.012)*(35) + (-0.0003)*(35^2) + (-0.694*1) + (1.935*1) + (0.112*1)))))
```
For the logic model, the predicted probability for a 35 year old female naturalized citizen with an college degree to be part of the labor force is 89.21%

```{r}
1/(1+(exp(-(0.707 + (0.012)*(35) + (-0.0003)*(35^2) + (-0.694*1) + (1.935*1) + (0.465*1)))))
```
For the logic model, the predicted probability for a 35 year old female non citizen with an college degree to be part of the labor force is 92.17%

For the probit model we can predict the probability of a 35 year old female naturalized citizen and non citizen with an college degree participates the labor force considering the independent variables as as AGE, AGE^2, FEMALE, EDUC_COLLEGE, CITIZENNaturalized citizen and CITIZENNot a citizen. 

```{r}
pnorm(0.707 + (0.012)*(35) + (-0.0003)*(35^2) + (-0.694*1) + (1.935*1) + (0.112*1))
```
For the probit model, the predicted probability for a 35 year old female naturilized citizen with an college degree to be part of the labor force is 98.27%

```{r}
pnorm(0.707 + (0.012)*(35) + (-0.0003)*(35^2) + (-0.694*1) + (1.935*1) + (0.465*1))
```
For the probit model, the predicted probability for a 35 year old female non citizen with an college degree to be part of the labor force is 99.32%

When comparing the probabilities calculated from the logit model 2, there is a 3 percentage point difference for women with college degrees as a naturalized citizen compared to women of the same parameters that are non citizens. 

When comparing the probabilities calculated from the probit model 2, there is a 1.05 percentage point difference for women with college degrees as a naturalized citizen compared to women of the same parameters that are non citizen. 

Lastly, when comparing the probabilities calculated from the probit and logit model 2, the difference is 9.06 percentage points for naturalized citizens and 7.15 percentage points for non citizens. The probit model shows a stronger predictability of determining whether a female either with naturalized citizen ornon citizen status with a college degree is part of the labor force.

```{r}
#modellin <- lm(LABFORCE ~ AGE + I(AGE^2) + female + AfAm + Asian + race_oth + Hispanic + educ_hs + educ_somecoll + educ_college + educ_advdeg + MARST +CITIZEN, data = dat_use2)
#summary(modellin)
```
Finally, when running the OLS model for the variables used in the probit and logit model 2, we encounter an error message since the dependent variable is binary and not numeric. Therefore, a logic and probic model is ideal to anaylze a non linear relationship for labor force and independent variables that are non numeric like citizen status or marital status.


#### Review of Research Articles

Reading below done by: Christopher Tinevra

Research Article 1 Title: "Work from Home After the COVID-19 Outbreak"

Authors: Alexander Bick, Adam Blandin, and Karel Mertens.

The research article emphasizes the recent changes to the workforce due the effect caused by the current global pandemic. Indeed, the research illustrates a shift from the regular workplace to remote work following the advice and mandates set by government officials. Furthermore, the results which were estimated from a sample size of 5,000 workers shows that majority of the workers can effectively shift and work from home. The econometric techniques used in the research include linear regression models where a correlation was formed in Figure 1 plotting a positive regression lines for the change of remote working by various industries. Additionally, multiple combine statistic tables further illustrate the change in commuting to work, changes in employment, and workers characteristics pre-pandemic and post-pandemic. In conclusion, the research provides potential evidence of the effect of the pandemic on the workforce by changing the workplace to a remote home setting for most workers depending on industry business and the possible job losses since the beginning of the outbreak. 

Reference: Bick, Alexander and Blandin, Adam and Mertens, Karel, Work from Home after the Covid-19 Outbreak (June, 2020). FRB of Dallas Working Paper No. 2017, Available at SSRN: https://ssrn.com/abstract=3638737 or http://dx.doi.org/10.24149/wp2017

Research Article 2 Title: "Who Telecommutes? Where is the Time Saved Spent?"

Authors: Harley Frazis

In this research paper, the growth of telecommuting is examined while taking into consideration ACS data that has illustrated a yearly increase in the number of workers that work from home in the last 15 years. Furthermore, the result estimates were gathered from the sample size of approximate 10,000 respondents with continues data responses ranging from the years 2003 to 2018 through the use of the American Time Use Survey. The methodology of the research paper includes two main approaches including simple comparison between the commuting and non-commuting workdays for the workers, while also comparing the time use or allocated by telecommuters to non-telecommuters. By using regression analysis, the paper demonstrates the further growth of telecommuting as well as how time is reallocated in comparison with time spend on commuting to work. In conclusion, the research paper provided some interesting findings showing the estimate of reallocated time for telecommuters is majorly spent on leisure while also illustrating the continue growth of telecommuting in recent years from the typical commuting to workplace jobs. 

Reference: Frazis, H. (2020). Who telecommutes? Where is the time saved spent?. BLS. Available at: https://www.bls.gov/osmr/research-papers/2020/ec200050.htm

Reading below done by: Nicole Kerrison

Research Article 1 Title: "Covid 19 mortality is Negatively Associated with Test Number and Government Effectiveness"

Authors: Li‑Lin Liang, Ching‑HungTseng , Hsiu J. Ho & Chun‑Ying Wu

This study aims to identify why the mortality rate of Covid-19 is varied across nations. While it is clear that there is a correlation with preexisting condition and age with mortality rate, other factors must be contributing to the variation in mortality rate across the world. Data was selected from the "Worldometer website" which list the coronavirus cases and mortality rates  for all countries and pulls data from multiple sources (WHO, US Center for Disease Control, and Computational Health Informatics at Harvard). Data was pulled from website and used as of June 13, 2020. Variable for Covid mortality was used and log transformed to make the data conform to a normal distribution and the regression analysis is ran using Stata software. A linear regression analysis was conducted across Covid-19 mortality rate and it was determined that its negatively associated with Covid-19 test number per 100 people, government effectiveness score, and number of hospital beds. The study also revealed that the Covid-19 mortality rate was positively associated with proportion of population aged 65 or older and transport infrastructure quality score.

Journal: Scientific Reports
www.nature.com/scietificreports

Research Article 2 Title: "An Ecological Study of Socioeconomic Predictors in Detection of COVID-19 Cases Across Neighborhoods in New York City"

Authors: Richard S Whittle and Ana Diaz-Artiles

This study aims to analyze the positive COVID-19 cases across the 5 boroughs of New York City since NYC became the epicenter of the virus in the United States. It seems that certain neighborhoods are experiencing more cases than others and this paper attempts to explain this by looking at socioeconomic differences in these boroughs. By using spatial modeling techniques and regression analysis, they could determine if key factors (Race, population density, income, and youth population) could predict clusters of positive tests.They collected data from the Zip Code Tabulation Areas (ZCTA) and used Bayesian Besag-York- Mollie (BYM) models where COVID-9 tests were the outcome. From there, they included a set of 11 demographics and parameters associated with the ZCTA in order to predict areas where COVID-19 prevalence would be. The study determined that there is a significant correlation between COVID-19 positive test rate and neighborhoods with a large percentage youth population, neighborhoods that are predominately Black, and low-income areas.

Journal: BMC Medicine
https://doi.org/10.1186/s12916-020-01731-6

